\documentclass[11pt,a4paper]{article}

% =========================================================
% EASY CONTROLS
% Change only these first when you want to tune the layout.
% =========================================================
\usepackage[
    a4paper,
    top=1.0in,
    bottom=1.0in,
    left=1.02in,
    right=1.02in
]{geometry} % Make left/right smaller to widen the text block

\newcommand{\TitleFont}{\fontsize{15.5}{17.5}\selectfont}
\newcommand{\AuthorFont}{\normalsize}
\newcommand{\AffilFont}{\small}
\newcommand{\AbstractFont}{\small}
\renewcommand{\baselinestretch}{1.04} % Slightly fuller text without looking loose

% =========================================================
% PACKAGES
% =========================================================
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{newtxtext,newtxmath} % Times-like research paper font
\usepackage{microtype}           % Better line breaking / more paper-like justification
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{array}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{titlesec}
\usepackage{caption}
\usepackage{setspace}
\usepackage{xcolor}

% =========================================================
% GENERAL FORMATTING
% =========================================================
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    urlcolor=blue,
    citecolor=blue
}

\setlength{\parindent}{1.2em}
\setlength{\parskip}{0pt}
\setlength{\emergencystretch}{1.5em} % Helps reduce overfull lines gracefully

\titleformat{\section}{\large\bfseries}{\thesection}{0.7em}{}
\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection}{0.7em}{}
\titlespacing*{\section}{0pt}{1.4ex plus 0.4ex minus 0.2ex}{0.7ex}
\titlespacing*{\subsection}{0pt}{1.0ex plus 0.3ex minus 0.2ex}{0.5ex}

\captionsetup{
    font=small,
    labelfont=bf
}

\setlist[itemize]{
    leftmargin=1.4em,
    itemsep=0.15em,
    topsep=0.25em
}

\setlist[enumerate]{
    leftmargin=1.6em,
    itemsep=0.15em,
    topsep=0.25em
}

% =========================================================
% TITLE
% =========================================================
\title{}
\author{}
\date{}

\begin{document}

\begin{center}
{\TitleFont\bfseries
DBSHIELD --- DataBase Systems Handling for Injection, Exploitation, Leakage, and DoS
\par}

\vspace{0.65em}

{\small
\begin{tabular}{c}
Arnav Priyadarshi (23CS30008) \quad
Ketan Suman (23CS30027) \quad
Krishna Prabu (23CS30028) \\[0.25em]
Rahul Chandrakant Patne (23CS10084) \quad
Shreeraj Kalbande (23CS30025) \\[0.35em]
\end{tabular}
\par}

\vspace{0.7em}
\end{center}

\section{Problem Statement}

Modern web applications depend heavily on databases to store and process sensitive information such as credentials, personal records, academic data, and administrative entries. Thus, it is important to study and understand how to construct secure databases. Ignoring this often leads to insecure design decisions such as unsafe query construction, weak validation of user-controlled input, insufficient server-side access checks, and inadequate controls against computational abuse. When such weaknesses exist in a database-backed system, the consequences can include unauthorized access, leakage of sensitive information, loss of service availability and more.

We present the \textbf{DBSHIELD} project, motivated by the need to demonstrate that these issues are not merely theoretical, but can arise in many places during development. Many users and developers underestimate how severely a database system can be affected when core security mechanisms are absent or improperly implemented. Instead of discussing these risks only in abstract terms, we aim to present through this project a direct comparison between two versions of the same application: an \textit{unshielded} system that is intentionally vulnerable for controlled demonstration, and a \textit{shielded} system that includes appropriate safeguards. This side-by-side design makes the role of security measures obvious, and shows how without them the system can be significantly harmed.

The application selected for this project is a mock student ERP system with three roles: \textit{student}, \textit{teacher}, and \textit{admin}. We selected this setting as it is something the intended audience can relate to easily, and reflects a realistic role-based application in which users authenticate, access restricted information, perform role-specific actions, and interact with shared database records. Such a system will naturally expose the central security requirements of authentication, data isolation and resilience. It therefore provides a practical environment in which database related vulnerabilities and their mitigations can be examined in a meaningful way.

We focus on three representative classes of attack. The first is \textit{SQL injection}, which targets the interaction between user input and backend query execution. When queries are not handled safely, SQL injection may allow login bypass, unauthorized reads, and unintended modifications to database contents. The second is \textit{denial of service}, studied here in the form of expensive application or database requests that can degrade responsiveness or exhaust resources. The third is \textit{authorization bypass}, where a user attempts to access resources or operations outside the limits of their intended role, particularly in systems that trust client-side verification such as cookies over server-side verification.

% The central aim of DBSHIELD is not to build an offensive toolset, but to illustrate the importance of defensive engineering in database-backed applications. By observing how the unshielded system behaves under controlled adversarial testing and comparing it with the hardened system, the project seeks to show that common safeguards can significantly reduce risk. In this sense, DBSHIELD is both a comparative security study and a practical demonstration of why secure database interaction, strict role enforcement, and protective backend design are essential in modern software systems.

\section{Methodology}

\subsection{Overall Study Design}

Two functionally similar versions of the same student ERP application are designed and studied:

\begin{itemize}
    \item \textbf{Unshielded system:} A deliberately vulnerable implementation in which protections are absent, to demonstrate the importance of defending against attacks.
    \item \textbf{Shielded system:} A hardened implementation that preserves the same functionality while incorporating defensive mechanisms to prevent attacks.
\end{itemize}

This paired design helps isolate the effect of security controls. Since both systems serve the same use case and operate on the same general model of data and user roles, differences in observed behavior can be attributed more clearly to the presence or absence of safeguards rather than to unrelated architectural changes.

\subsection{System Under Study}

The application modeled in this project is a student ERP system with three user roles:

\begin{itemize}
    \item \textbf{Student:} Accesses personal details, academic information, and self-related records. Can enroll in courses, deregister from courses, but not view or interact with other students.
    \item \textbf{Teacher:} Accesses and updates academic information related to students and courses. Can create courses, modify contents, but cannot modify other teacher's courses.
    \item \textbf{Admin:} manages users, records, and privileged system-level operations.
\end{itemize}

This role structure is intentionally chosen because it captures a realistic multi-user scenario in which different permissions must be enforced over the same database. It provides a suitable basis for studying authentication flows, role-based restrictions, protected data access, and misuse of sensitive operations.

\subsection{Technology Stack}

The system uses a web-application stack composed of the following components:

\begin{itemize}
    \item \textbf{Frontend:} React is used to build the user-facing interface, including login pages, dashboards, navigation, and role-specific views.
    \item \textbf{Backend:} FastAPI is used to implement API endpoints, application logic, authentication handling, request processing, and access control logic.
    \item \textbf{Database:} PostgreSQL and SQLite is used as the relational data store for user accounts, role information, academic records, and related entities.
\end{itemize}

\subsection{Unshielded and Shielded Implementations}

In both the cases, the FastAPI backend is responsible for retrieving from the database, perform logical operations, checking and communicating the result to the frontend.

In the \textit{Unshielded Implementation}, the raw SQL is directly fed, without any additional checking, to the database to fetch data. This can cause issues such as SQL Injection, DoS and Authorization Bypass, which are discussed later. The result similarly is directly fed to the user, without verification.

In the \textit{Shielded Implementation}, the input from the frontend is processed, checked for potential SQL Injection attacks, checked for DoS attacks, and is only then processed. This avoids issues that used to arise in the unshielded implementation.

\subsection{SQL Injection Study}

The SQL injection study now covers both authentication and several ERP data-handling endpoints. The backend preserves an intentionally vulnerable implementation in its original service layer, while the hardened behavior is implemented in a separate module under \texttt{sql\_injection\_prevention}. A single runtime flag, \texttt{ENABLE\_SQLI\_PROTECTION}, determines which handler is used for each request. This paired design is useful for controlled evaluation because the application routes, payload formats, and user-visible functionality remain unchanged, while only the query-construction strategy differs between the two modes.

In the unshielded mode, multiple operations continue to build SQL directly from user-controlled values. These include login, course search, grade and course lookups based on usernames, enrollment and grading workflows, administrative creation and deletion actions, and an especially sensitive admin query endpoint. In the shielded mode, the same operations are routed to secure counterparts that use parameterized queries for user-supplied values. For the admin query interface, which accepts free-form SQL text by design, the secure path does not attempt to parameterize arbitrary statements; instead, it restricts execution to a single read-only \texttt{SELECT} or \texttt{WITH} query and blocks mutating or stacked statements.

To evaluate this design, we use a controlled set of payloads against both the login path and representative ERP endpoints. The test corpus includes authentication-bypass inputs based on comments and tautologies, \texttt{UNION}-style payloads intended to forge result rows or enumerate schema information, and username or course-code payloads intended to alter selection or deletion behavior in routine ERP operations. Additional tests also check whether values that look like SQL fragments, such as a malicious-looking grade string or username, are stored and handled as literal data in the secure mode rather than being executed as part of the query.

The experiments are executed through the project test harness and through the application toggle that switches between vulnerable and secure handlers. For each case, the main observation is whether the supplied payload changes query semantics, broadens the result set, causes unintended data modification, or bypasses the intended restriction on the admin action endpoint. The comparison therefore measures not only whether login injection is prevented, but also whether the same defensive strategy generalizes across the broader ERP workflow.

\subsection{Denial of Service Study}

The second attack category focuses on denial of service in the context of application and database resource abuse. In this project, denial of service is not treated as a raw network flooding exercise, but as a study of whether expensive or poorly bounded requests can degrade responsiveness, consume disproportionate resources, or make the system unstable. The unshielded system is examined using controlled requests that are expected to trigger heavy backend or database work.

The shielded system is then evaluated under the same workload patterns. The expectation is that defensive mechanisms such as better request filtering, bounded execution, query restrictions, or rate-aware handling will reduce performance degradation and improve service stability. The comparison emphasizes availability and resilience rather than only whether a single request succeeds.

\subsection{Authorization Bypass Study}

The third attack category studies authorization bypass. Because the ERP has clearly separated user roles, it is important to evaluate whether a user can access functions or records beyond their intended permissions. This includes scenarios such as requesting pages intended for another role, directly invoking restricted backend endpoints, or attempting to retrieve or modify records that should not be accessible under the current identity.

The unshielded system is used to study whether role separation is enforced weakly or only on the client side. The shielded system is then used to verify whether access control is consistently enforced on the server for every protected action. The comparison highlights the difference between superficial interface-based restrictions and proper backend authorization.

\subsection{Evaluation Criteria}

For each attack category, the final comparison between the unshielded and shielded systems is organized using the following criteria:

\begin{itemize}
    \item \textbf{Attack success:} whether the intended misuse achieves its goal.
    \item \textbf{Impact scope:} what data, role, or service is affected.
    \item \textbf{Response behavior:} whether the system leaks information, hangs, returns unsafe errors, or fails in a controlled way.
    \item \textbf{Resilience:} whether the application remains stable and recoverable during and after testing.
    \item \textbf{Mitigation effectiveness:} whether the shielded system prevents or significantly limits the observed issue.
\end{itemize}

\section{Planned Remaining Structure}

The report is intended to expand into a compact ten-page document. After the problem statement and methodology, the remaining sections can be developed in the following order:

\begin{itemize}
    \item a section describing the system architecture and technology choices in more detail,
    \item one section each for SQL injection, denial of service, and authorization bypass,
    \item a results section comparing the unshielded and shielded implementations,
    \item a references section.
\end{itemize}

Each attack section can later describe the attack goal, the part of the ERP being tested, the controlled test cases used, and the expected difference between the vulnerable and hardened versions of the system.

\section{Results}

\subsection{SQL Injection Results}

The updated SQL injection experiments show a consistent separation between the vulnerable ERP service layer and the shielded handlers selected through \texttt{ENABLE\_SQLI\_PROTECTION}. On the login path, the previously documented SQLite-compatible payloads continued to succeed against the unshielded implementation: comment-based and tautology payloads enabled authentication bypass, \texttt{UNION}-based payloads could fabricate or expose rows, and blind inference payloads could still alter query logic. This confirms that the vulnerable path remains suitable for controlled demonstration.

The broader ERP tests show the same pattern. In the unshielded mode, username-based injection payloads could broaden student-grade and student-course lookups, and course-search input could alter selection behavior. Administrative actions that interpolate course codes or usernames are likewise exposed in the vulnerable path because the query text is still assembled directly from request data. In contrast, the shielded handlers treat these same payloads as literal values. Search and lookup functions return no unauthorized data, delete operations do not affect unintended rows, enrollment requests using injected usernames fail normally, and malicious-looking values can be stored as plain data without changing query structure.

The hardened admin action endpoint is a particularly important result. Since this interface is meant to accept SQL text, parameterization alone is not sufficient. The secure implementation therefore adds a structural restriction: only a single read-only \texttt{SELECT} or \texttt{WITH} statement is accepted. As a result, mutating statements such as \texttt{DROP TABLE}, \texttt{UPDATE}, or stacked multi-statement payloads are rejected, while legitimate read-only queries continue to work. This makes the defense more realistic for a free-form admin query feature, where the goal is controlled capability rather than complete removal of functionality.

\begin{table}[h]
\centering
\caption{Observed outcome of SQL injection tests after adding toggle-based secure handlers}
\begin{tabular}{p{0.31\linewidth} p{0.28\linewidth} p{0.28\linewidth}}
\toprule
\textbf{Payload category} & \textbf{Unshielded system} & \textbf{Shielded system} \\
\midrule
Comment and tautology login bypass & Successful unauthorized login & Blocked; login fails normally \\
\texttt{UNION}-based data extraction & User data and query-controlled rows returned & Blocked; no data disclosure \\
\texttt{UNION}-based role forgery and schema manipulation & Forged or metadata-oriented rows can be returned & Blocked; payload treated as data \\
Injected usernames in student lookups & Result set can be broadened beyond intended user & Blocked; no unauthorized rows returned \\
Injected course codes in admin delete/search paths & Query behavior can be altered & Payload treated literally; intended row scope preserved \\
Injected values in enroll/grade/create operations & May alter query semantics in vulnerable flow & Stored or rejected as plain data without execution \\
Mutating or stacked admin SQL & Can execute directly on vulnerable path & Rejected unless it is a single read-only query \\
\bottomrule
\end{tabular}
\end{table}

Overall, the results show that the project has moved from a single protected login demonstration to a broader SQL injection defense layer covering multiple ERP operations. The most important outcome is that the secure path does not rely only on ad hoc blacklist filtering for ordinary endpoints; instead, it uses parameterized queries so that attacker-controlled input no longer changes SQL semantics. Where arbitrary SQL input is unavoidable, as in the admin action feature, the secure path applies explicit read-only restrictions. Together, these changes substantially reduce the number of practical SQL injection paths while preserving the intentionally vulnerable path for demonstration and comparison.

\section{References Placeholder}

A later version of the report can include references to relevant documentation and defensive security material related to the stack and attack categories used in the project.

\begin{thebibliography}{9}

\bibitem{owasp-top10}
OWASP Foundation.
\textit{OWASP Top 10}.
Available online.

\bibitem{owasp-sqli}
OWASP Foundation.
\textit{SQL Injection Prevention Cheat Sheet}.
Available online.

\bibitem{postgres}
PostgreSQL Global Development Group.
\textit{PostgreSQL Documentation}.
Available online.

\bibitem{fastapi}
FastAPI.
\textit{FastAPI Documentation}.
Available online.

\bibitem{react}
React.
\textit{React Documentation}.
Available online.

\end{thebibliography}

\end{document}
